A self-consistency check for unitary propagation of Hawking quanta

### 2016-11-02 Draft_BH_shells.tex

\begin{abstract}
The black hole information paradox presumes that quantum field theory in curved spacetime can provide unitary propagations from near-horizon modes to asymptotic Hawking quanta. 
Instead of invoking conjectural quantum-gravity effects to modify such assumption, we propose a self-consistency check.
We establish an analogy to Feynman's analysis of a double-slit experiment. 
Feynman showed that a unitary propagation of the interfering particles, namely ignoring the entanglement with the double-slit, becomes an arbitrarily reliable assumption when the screen to project the interference pattern goes to infinitely far away.
We argue for an analogous self-consistency check for quantum field theory in curved spacetime.
We apply it to the propagation of Hawking quanta, testing whether ignoring the entanglement with the geometry also becomes arbitrarily reliable in the limit of a large black hole. 
We present curious results to suggest a negative answer, and we discuss how this loss of na\"ive unitarity in QFT is related to the soft-hair-memory effect.
\end{abstract}

@@

The black hole information paradox \cite{Haw76a} was sharpened to show a self-inconsistency among the following three widely-believed statements \cite{AMPS}
\footnote{Different physicists may believe in some of those more strongly than the others, but very few would have argued against any single statement without the explicitly conflict in the information paradox.}:
\begin{itemize}
\item {\bf Unitary evaporation:} A black hole will totally evaporate; the formation and evaporation of a black hole is described by an asymptotic observer as a unitary $S$-matrix, whose size is the exponential of the black hole's Bekenstein entropy.
\item {\bf General relativity:} The collapse-Schwarzschid geometry given by general relativity is a valid description to the spacetime everywhere except for near the singularity.
Therefore, there is no drama while crossing the horizon.
\item {\bf Local quantum field theory:} Away from the singularity, we can apply local quantum field theory (QFT) which describes microscopic steps of evaporation. 
For large black holes, it manifests itself as unitary propagation of every near-horizon mode into every asymptotic Hawking quanta.
\end{itemize}
Basically, Statement Three demands that a near horizon mode is related to an asymptotic Hawking quanta by a unitary transformation, therefore carrying the same qubit of information. 
However, Statement One and Two demand this unique qubit to be maximally entangled with two distinct objects, violating monogamy. 
This conflict has inspired many different proposals to modify one of the above three statements. 
For example, information loss or remnant \cite{Bek94} modifies Statement One; 
firewall \cite{BraPir09,AMPS} or ER=EPR \cite{MalSus13} modifies Statement Two;
various proposals of nonlocal effects near the horizon \cite{Gid12,DodSil15,OsuPag16}, causal patch complementarity \cite{HuiYan13,IlgYan13,LowTho14}, or computational complementarity \cite{HH} modify Statement Three.
\footnote{Various versions of complementarity basically claim that QFT breaks down if one applies it compare quantities which are in-principle not measurable (or difficult to measure) by the same observer.}

In this paper, we will try to provide a stronger motivation to modify Statement Three. 
In fact, we will argue that it is already problematic on its own. 
In other words, {\bf the application of local QFT to a unitary propagation of Hawking quantum is not self-consistent. }

@@

One proposal along this line of thoughts was discussed in \cite{NomVar12}. 
The idea was that the black hole acquires a large uncertainty during the evaporation process. 
If we treat the superposition of different classical geometries as a quantum system, it may qualify as the hidden system that is responsible for the non-unitary propagation of Hawking quanta. 
Unfortunately, the early version of such proposal can be defeated by counting entropy.
In order for such non-unitary process to restore the asymptotic $S$-matrix, the hidden system must store the required information, which is comparable to the amount of microscopic information carried by the black hole.
However, the black-hole-no-hair theorem states that two black holes with identical macroscopic parameters (mass, momentum, angular momentum, charge) are isometric to each other. 
This limits the number of different classical geometries to be parametrized by macroscopic parameters only.
As a result, the Hilbert space of superposed classical geometries has far fewer dimensions than what is necessary to carry the required amount of information.
\footnote{We thank Raphael Bousso for private communication.}
Thus, even if a propagating Hawking quantum does interact with this hidden system, it cannot reproduce a unitary $S$-matrix; therefore such modification of Statement Three alone is insufficient to resolve the paradox.

@@

In Sec.\ref{sec-BlackHole}, we apply the self-consistency check to the propagation of Hawking quanta in a close analogy to Feynman's analysis. 
The classical geometry plays the role of the double-slit, a background that can potentially entangle with the Hawking quanta wavefunction and ruin its subsystem unitarity. 
We then assume that the unknown quantum nature of geometry can be parametrize by a classical uncertainty with an unknown but fixed, gauge-invariant size. 
A classical wave plays the role of the interference pattern. 
When even a classical wave decoheres due to the uncertainty of geometry, there is no reason to believed that the quantized version of such wave has unitarity.
The common expectation is that in the limit of a large black hole, gravitational effects outside the horizon are arbitrarily weak, thus local QFT should be arbitrarily trustworthy.
By analogy to Feynman's analysis, we should expect that with a fixed uncertainty in the geometry, any effect on the classical wave should drop to zero when we take the large black hole limit.
Interestingly, what we found seems to be the opposite. 
We show that with a fixed geometric uncertainty, a classical wave suffers an {\it arbitrarily large} correction in the large black hole limit.
By analogy to Feynman's analysis, this is a strong evidence that the apparent unitarity of QFT in this case is not a valid assumption.

@@

\begin{figure}[h!]
\begin{center}.
\includegraphics[scale = 0.5]{DSe.pdf}
\caption{Schematic of the double slit experiment. The diagram shows two different scenarios. The top figure shows the behaviour one would expect of classical particles going through one slit or the other. The bottom figure shows the behaviour of classical waves creating an interference pattern. Screen one shows the signal that one would see in each scenario (i.e classical particles or waves) if only one slit was open. If slit 1 was open we would see $P_1$ and see on. Screen 2 shows the signal we see if both the silts are open in each scenario.}
\label{fig-doubleslit}
\end{center}
\end{figure}

@@

If we go back to the basic of quantum field theory, there is a natural description of this new rule. 
As shown in Fig.\ref{fig-QFT}, QFT starts from solving a classical mode function, then quantizing its amplitude into a quantum state. 
Any method to measure such quantum state assumes the knowledge of the classical mode function. 
If the geometric uncertainty leads to a significant uncertainty in the classical mode function, then there is no practical way to reliably measure its quantum state. 
We advocate that this is not only a practicality about measuring the quantum state, but it directly tells us that the quantum state loses its unitarity. 
Just like in Feynman's example where the uncertainty of the double-slit guarantees its entanglement with the particles, whatever effect that leads to the geometric uncertainty here must entangle with the quantum state of this mode. 
Without a theory of quantum geometry, we cannot describe how that happens. 
But through a classical calculation, we can determine whether it has happened or not.

@@

While such generalization to QFT seems straightforward in theory, there are a few subtleties in practice.
In the double-slit experiment, the position uncertainty of the double-slit was the only obvious uncertainty of the background that we need to keep track of.
It is also the reasonable thing to hold fixed while moving the projection screen away.
In QFT, the background is the entire geometry, and there are infinitely many ways a geometry can be uncertain.
For a full analogy to Feynman's analysis, not only we need to parametrize those uncertainties, we also need to choose the appropriate combination to hold fixed as we take a similar limit.
At this moment, we do not have a general formalism to setup such self-consistency check that can be applied to any general geometry. 
In Sec.\ref{sec-BlackHole}, we propose a specific way to apply such self-consistency check to the propagation of Hawking quanta.
In addition to gaining insights into the information paradox, we hope to learn some lessons to inspire a more general formalism of such self-consistency check.

@@

\begin{itemize}
\item {\bf Classical coherence by tracking geodesics.} The classical mode of a massless field basically has its peaks and nodes following null geodesics.
Thus, instead of actually solving the mode function, we can perform a much simpler calculation to track geodesics.
We start with two out-going null geodesics with the separation of one period of such mode near the horizon, and then we calculate the change in their separations at spatial infinity.
As this change increases, one should gradually lose faith in the unitary propagation of such mode, since it becomes increasingly difficult to experimentally verify.
\item {\bf The background and the uncertainty.} The classical background in this case is the Schwarzschild geometry, which can be written as the metric $g_{\mu\nu}$. 
A classical uncertainty can be expressed as deviation from such metric, $\Delta g_{\mu\nu}$. 
There are two major challenges here.
\begin{enumerate}
\item There are many more variables than the double-slit experiment. 
We consider only one particular form of $\Delta g_{\mu\nu}$ in this paper---back-reactions from the presence of extra matter of zero total energy. 
It is simple to calculate and reflects the physcial intuition of vacuum pair fluctuation.
\item The value of $\Delta g_{\mu\nu}$ is gauge dependent. 
By relating it to the presence of extra matter, we can express it as a gauge-invariant local quantity, which is the natural thing to hold fixed in our analysis.
\end{enumerate}
Note that if we were reporting a positive result, that such uncertainty upholds unitarity, then one should question the validaty of learning a general lesson from a special example. However, we will show that this particular form of $\Delta g_{\mu\nu}$ is already sufficient to question unitarity, and we did not choose this form to specifically do that. Thus our limited analysis is sufficient to raise a reasonable doubt to the classical-background approximation.
\item {\bf Renormalization.} It is well-known that na\"ive applications of QFT often suffer from UV/IR divergences.
In our case, we will see that even in Minkowski space, geometric uncertainty already leads to a finite change between the two null geodesics.
We will simply ``renormalize'' that away by saying that if the black hole case leads to a similar change, it should not be taken as a serious problem.
Our surprising result is that in the black hole case, we get an infinitely larger change, which is difficult to blame on the usual divergence of QFT.
\end{itemize}

More concretely, our setup is shown in Fig.\ref{fig-setup}. 
We start by choosing two points at some $r_0\gg M$ with a $\delta t_0\sim M$ coordinate time separation between them.
This represents one wavelength of the mode function of an asymptotic Hawking quantum.
We then back-track two null geodesics from these two points back to two points near the horizon, at $r_e$ with $(r_e-2M)\ll 2M$.
Now their separation represents one wavelength of a near-horizon mode, which would have propagated to the asymptotic mode if there were no geometric uncertainty. 
We then introduce introduce a geometric uncertainty as two shells of zero total ADM mass. 
Therefore, both the near horizon and the asymptotic regions are not affected by such uncertainty.
Only the process of propagation is affected.
With the same two starting points near the horizon. we calculate the separation $\bar{\delta t_0}$ of the two null geodesics when they arrive at $r_0$, and compare it with $\delta t_0$.

\begin{figure}[tb]
\begin{center}
\includegraphics[scale = 0.6]{Propertime.pdf}
\caption{The two parts of this diagrams describe two different scenarios. The trajectories of the photon are the lines in orange. The trajectory of the observer and the position from which the photons are emitted are represented by solid blue lines. The bottom part is a general case of two Hawking photons coming from a distance $r_A$ from the black hole of mass $M$ and arriving at an observer who is at a distance of $r_0$. The top part shows two Hawking photons coming from the same distance $r_A$ from a black hole of mass $M$. Instead of the photons freely propagating through to an observer at $r_0$, they have to cross two shells, represented by blue dotted line, of equal and opposite mass $\Delta M, -\Delta M$ at $r_{1}, r_{2}$ respectively. The regions $I$ represent a metric with mass $M$ and region $II$ represents a metric with mass $M+\Delta M$.}
\label{fig-setup}
\end{center}
\end{figure}

@@

After some formal calculations in Appendix \ref{Dtime}, we arrive at this simple expression:
\begin{equation}
\frac{\delta t_0 - \bar{\delta t_0}}{\delta t_0} = 
\left(\sqrt{\frac{r_2-2M}{r_1-2M}}\sqrt{\frac{r_1-2(M+\Delta M)}{r_2-2(M+\Delta M)}}-1\right)~,
\label{eq-result}
\end{equation}
where $r_1$ is the location of the inner shell, $r_2$ is the location of the outer shell, and $\pm\Delta M$ are their individual ADM masses.

@@

In the entire calculation, we will stay in the regime that $r_1, 2M, (r_1-2M) \gg \Delta M, \Delta r$, where $\Delta r \equiv(r_2-r_1)$. 
This guarantees that the geometric fluctuation is small, as it only affects a small region $\Delta r$, and the change in the metric $\Delta g_{\mu\nu}/g_{\mu\nu}$ is small in the affected region.
Under this assumption, we can expand Eq (\ref{eq-result}) to get 
\begin{equation}
	\frac{\delta t_0 - \bar{ \delta t_0}}{\delta t_0} = \frac{\Delta M \Delta r}{(r_1-2M)^2} = \frac{\Delta M \Delta r}{r_1^2} g_{tt}^{-2}~.
		\label{lim-result}
\end{equation}
Here $g_{tt}=(1-2M/r_1)$ is the radial component of the metric in the background geometry, evaluated at the location where the null rays pass through the geometric uncertainty.

It is more enlightening to write this expression in the local physical quantities (see Appendix \ref{finalEq} for derivations).  Let $\sigma$ be the energy density of the shell in its rest frame, and $\Delta l$ be the physical distance between the shells in that frame, we get
\begin{equation}
	\frac{\delta t_0 - \bar{ \delta t_0}}{\delta t_0} = 4 \pi \sigma \Delta l g_{tt}^{-1}~.
	\label{eq-inv}
\end{equation}
Here we can see that the effect is due to the local energy density which the null rays pass through.
Also, the combination $(\sigma\Delta l)$ is invariant under radial boosts, thus Eq.~(\ref{eq-inv}) does not care about whether the shells are actually at rest, nor in which frame we calculate it.

@@

\begin{equation}
\frac{\delta t_0 - \bar{ \delta t_0}}{\delta t_0}  \begin{cases}
= 4 \pi \sigma \Delta l~	 &\ \ \ \ \ (M = 0)\\
\approx 4\pi \sigma \Delta l \left( 1 + \frac{2M}{l} \right) & \ \ \ \ \ (M \ll l)	\\
\approx 4 \pi  \sigma \Delta l \frac{16M^2}{l^2} & \ \ \ \ \ (M \gg l).
\end{cases}	\label{fin-result}
\end{equation}

@@

Classically, soft hairs manifest as memory effects that alters the distances between geodesics, which is very similar to the effect of geometric uncertainty as we showed in this paper.
Two neighbouring null rays simply ``remember'' the geometric uncertainty they passed through.
Na\"ively speaking, a propagating mode of a massless field should remember a similar effect. 
Based on this picture, one may try to quantize the soft hairs and build a toy model that involves quantum interaction, therefore entanglement, between the soft hairs and the usual QFT modes.
That would provide a complete picture that the loss of subsystem unitarity is really just entanglement with another subsystem which was originally (and illegitimately, as we argued) assumed as the classical background.

@@

\begin{equation}
	ds^2 = - \left( 1 -\frac{2M}{r} \right) dt^2 + \left( 1 - \frac{2M}{r} \right)^{-1} dr^2 + r^2 d \Omega_2^2	\label{1}
\end{equation}

where we are working with units in which $G = c = 1$. We know there will be a flux of particles appearing from the horizon of the black hole \cite{Haw74}. Take some coordinate time interval, $\delta \bar{t_e}$, (which will have some corresponding proper time interval $\delta \bar{\tau_e}$) between two consecutive particles being emitted from the same position above the horizon of the black hole. An observer far away will see these particles arrive with some coordinate time interval $\delta \bar{t_0}$. We can follow the trajectory of the particles (assuming they are relativistic) from the position they are emitted, $r_e$, to the observer at $r_0$ by setting $ds = 0$ in Eq (\ref{1}). By confining the motion to be only in the radial direction, i.e $d \Omega_2 = 0$, we get the following geodesic equation

\begin{equation}
	\bar{t_e}^{(1)} - \bar{t_0}^{(1)} = \bar{r_0} - \bar{r_e} + 2M \ln \left( \frac{\bar{r_0} - 2M}{\bar{r_e} - 2M}  \right)	\label{3}
\end{equation}

@@

\begin{equation}
	\bar{t_e}^{(2)} - \bar{t_0}^{(2)} = \bar{r_0} - \bar{r_e} + 2M \ln \left( \frac{\bar{r_0} - 2M}{\bar{r_e} - 2M}  \right).	\label{4}
\end{equation}

@@

Now we introduce two shells of matter with equal and opposite ADM mass, $\Delta M$ and $-\Delta M$ where $|\Delta M|<<M$, at a distance of $r_1$ and $r_2$ respectively as shown in the top part of figure \ref{fig-setup}. We model these shells as infinitesimally thin and therefore use the Israel Junction Conditions (IJC) \cite{Isr66} with a delta function shell to analyze the effects of the shells. In this case the particle's move through two different spacetime regions. Region $I$ is defined by the black hole of mass $M$ and region $II$ is defined by a Schwarzschild metric of mass $M+\Delta M$ as shown in the top part of figure \ref{fig-setup}.
\\
\\
As explained in the previous section, the proper time interval between two particles emitted at $r_e$, $\delta t_e$, is the same as the coordinate time interval at $r_1$, $\delta t_1$. Now we need to find the corresponding coordinate time interval in region $II$, $\delta \hat{t}_1$ (in general we will use hats on top of quantities that are evaluated in region $II$). To do this we simply note that the corresponding proper time at $r_1$, $\delta \tau_1$ must be the same in both regions of spacetime - this is a simple consequence of the IJC. The coordinate time intervals at $r_1$, in the two different regions, are therefore related by

\begin{equation}
	\delta t_1 = \delta t_0 = \left( \frac{1 - \frac{2(M + \Delta M)}{r_1}}{1 - \frac{2M}{r_1}} \right)^\frac{1}{2} \delta \hat{t}_1.
\end{equation}

We can propagate the particles through region $II$ from $r_1$ to $r_2$ and we know $\delta \hat{t}_1 = \delta \hat{t}_2$. From the fact that the proper time corresponding to the coordinate time intervals will be the same, we can relate $\delta \hat{t}_2$ back to $\delta t_2$ in region $I$,

\begin{equation}
	\delta \hat{t}_2 = \delta \hat{t}_1= \left( \frac{1 - \frac{2M}{r_2}}{1 - \frac{2(M + \Delta M)}{r_2}} \right) \delta t_2.	\label{7}
\end{equation}

>>>>

\begin{abstract}
The black hole information paradox presumes that quantum field theory in curved spacetime can provide unitary propagations from near-horizon modes to asymptotic Hawking quanta. 
Instead of invoking conjectural quantum-gravity effects to modify such an assumption, we propose a self-consistency check.
We establish an analogy to Feynman's analysis of a double-slit experiment. 
Feynman showed that a unitary propagation of the interfering particles, namely ignoring the entanglement with the double-slit, becomes an arbitrarily reliable assumption when the screen to project the interference pattern goes to infinitely far away.
We argue for an analogous self-consistency check for quantum field theory in curved spacetime.
We apply it to the propagation of Hawking quanta, testing whether ignoring the entanglement with the geometry also becomes arbitrarily reliable in the limit of a large black hole. 
We present curious results to suggest a negative answer, and we discuss how this loss of na\"ive unitarity in QFT is related to the soft-hair-memory effect.
\end{abstract}

@@

The black hole information paradox \cite{Haw76a} was sharpened to show a self-inconsistency among the following three widely-believed statements \cite{AMPS}\footnote{Different physicists may believe in some of those more strongly than the others, but very few would have argued against any single statement without the explicit conflict in the information paradox.}:
\begin{itemize}
\item {\bf Unitary evaporation:} A black hole will totally evaporate; the formation and evaporation of a black hole is described by an asymptotic observer by a unitary $S$-matrix, whose size is the exponential of the black hole's Bekenstein entropy.
\item {\bf General relativity:} The collapse-Schwarzschid geometry given by general relativity is a valid description of the spacetime everywhere except for near the singularity.
Therefore, there is no drama while crossing the horizon.
\item {\bf Local quantum field theory:} Away from the singularity, we can apply local quantum field theory (QFT) which describes microscopic steps of evaporation. 
For large black holes, it manifests itself as unitary propagation of every near-horizon mode into every asymptotic Hawking quanta.
\end{itemize}
Basically, Statement Three demands that a near horizon mode is related to an asymptotic Hawking quanta by a unitary transformation, therefore carrying the same qubit of information. 
However, Statement One and Two demand this unique qubit to be maximally entangled with two distinct objects, violating monogamy. 
This conflict has inspired many different proposals to modify one of the above three statements. 
For example, information loss or remnant \cite{Bek94} modifies Statement One; 
firewall \cite{BraPir09,AMPS} or ER=EPR \cite{MalSus13} modifies Statement Two;
various proposals of nonlocal effects near the horizon \cite{Gid12,DodSil15,OsuPag16}, causal patch complementarity \cite{HuiYan13,IlgYan13,LowTho14}, or computational complementarity \cite{HH} modify Statement Three.\footnote{Various versions of complementarity basically claim that QFT breaks down if one applies it compare quantities which are in-principle not measurable (or difficult to measure) by the same observer.}

In this paper, we will try to provide a stronger motivation to modify Statement Three. 
In fact, we will argue that it is already problematic on its own. 
In other words, {\bf the application of local QFT to unitary propagation of Hawking quanta is not self-consistent. }

@@

One proposal along this line of thoughts was discussed in \cite{NomVar12}. 
The idea was that the black hole acquires a large uncertainty during the evaporation process. 
If we treat the superposition of different classical geometries as a quantum system, it may qualify as the hidden system that is responsible for the non-unitary propagation of Hawking quanta. 
Unfortunately, the early version of such proposal can be defeated by counting entropy.
In order for such non-unitary process to restore the asymptotic $S$-matrix, the hidden system must store the required information, which is comparable to the amount of microscopic information carried by the black hole.
However, the black-hole-no-hair theorem states that two black holes with identical macroscopic parameters (mass, momentum, angular momentum, charge) are isometric to each other. 
This limits the number of different classical geometries to be parametrized by macroscopic parameters only.
As a result, the Hilbert space of superposed classical geometries has far fewer dimensions than what is necessary to carry the required amount of information.\footnote{We thank Raphael Bousso for private communication.}
Thus, even if a propagating Hawking quantum does interact with this hidden system, it cannot reproduce a unitary $S$-matrix; therefore such modification of Statement Three alone is insufficient to resolve the paradox.

@@

In Sec.\ref{sec-BlackHole}, we apply the self-consistency check to the propagation of Hawking quanta in a close analogy to Feynman's analysis. 
The classical geometry plays the role of the double-slit, a background that can potentially entangle with the Hawking quanta wavefunction and ruin its subsystem unitarity. 
We then assume that the unknown quantum nature of geometry can be parametrize by a classical uncertainty with an unknown but fixed, gauge-invariant size. 
A classical wave plays the role of the interference pattern. 
When even a classical wave decoheres due to the uncertainty of geometry, there is no reason to believed that the quantized version of such wave has unitarity.
The common expectation is that in the limit of a large black hole, gravitational effects outside the horizon are arbitrarily weak, thus local QFT should be arbitrarily trustworthy.
By analogy to Feynman's analysis, we should expect that with a fixed uncertainty in the geometry, any effect on the classical wave should drop to zero when we take the large black hole limit.
Interestingly, what we found seems to be the opposite. 
We show that with a fixed geometric uncertainty, a classical wave suffers an {\it arbitrarily large} correction in the large black hole limit.
By analogy to Feynman's analysis, this is strong evidence that the apparent unitarity of QFT in this case is not a valid assumption.

@@

\begin{figure}[h!]
\begin{center}.
\includegraphics[scale = 0.5]{DSe.pdf}
\caption{Schematic of the double slit experiment. The diagram shows two different scenarios. The top figure shows the behaviour one would expect of classical particles going through one slit or the other. The bottom figure shows the behaviour of classical waves creating an interference pattern. Screen one shows the signal that one would see in each scenario (i.e classical particles or waves) if only one slit was open. If slit 1 was open we would see $P_1$ and so on. Screen 2 shows the signal we see if both the silts are open in each scenario.}
\label{fig-doubleslit}
\end{center}
\end{figure}

@@

If we go back to the basics of quantum field theory, there is a natural description of this new rule. 
As shown in Fig.\ref{fig-QFT}, QFT starts from solving a classical mode function, then quantizing its amplitude into a quantum state. 
Any method to measure such quantum state assumes the knowledge of the classical mode function. 
If the geometric uncertainty leads to a significant uncertainty in the classical mode function, then there is no practical way to reliably measure its quantum state. 
We advocate that this is not only a practicality about measuring the quantum state, but it directly tells us that the quantum state loses its unitarity. 
Just like in Feynman's example where the uncertainty of the double-slit guarantees its entanglement with the particles, whatever effect that leads to the geometric uncertainty here must entangle with the quantum state of this mode. 
Without a theory of quantum geometry, we cannot describe how that happens. 
But through a classical calculation, we can determine whether it has happened or not.

@@

While such generalization to QFT seems straightforward in theory, there are a few subtleties in practice.
In the double-slit experiment, the position uncertainty of the double-slit was the only obvious uncertainty of the background that we need to keep track of.
It is also the reasonable thing to hold fixed while moving the projection screen away.
In QFT, the background is the entire geometry, and there are infinitely many ways a geometry can be uncertain.
For a full analogy to Feynman's analysis, not only do we need to parametrize those uncertainties, we also need to choose the appropriate combination to hold fixed as we take a similar limit.
At this moment, we do not have a general formalism to setup such self-consistency check that can be applied to any general geometry. 
In Sec.\ref{sec-BlackHole}, we propose a specific way to apply such self-consistency check to the propagation of Hawking quanta.
In addition to gaining insights into the information paradox, we hope to learn some lessons to inspire a more general formalism of such self-consistency check.

@@

\begin{itemize}
\item {\bf Classical coherence by tracking geodesics.} The classical mode of a massless field basically has its peaks and nodes following null geodesics.
Thus, instead of actually solving the mode function, we can perform a much simpler calculation to track geodesics.
We start with two out-going null geodesics with the separation of one period of such mode near the horizon, and then we calculate the change in their separations at spatial infinity.
As this change increases, one should gradually lose faith in the unitary propagation of such mode, since it becomes increasingly difficult to experimentally verify.
\item {\bf The background and the uncertainty.} The classical background in this case is the Schwarzschild geometry, which can be written as the metric $g_{\mu\nu}$. 
A classical uncertainty can be expressed as deviation from such metric, $\Delta g_{\mu\nu}$. 
There are two major challenges here.
\begin{enumerate}
\item There are many more variables than the double-slit experiment. 
We consider only one particular form of $\Delta g_{\mu\nu}$ in this paper---back-reactions from the presence of extra matter of zero total energy. 
It is simple to calculate and reflects the physical intuition of vacuum pair fluctuation.
\item The value of $\Delta g_{\mu\nu}$ is gauge dependent. 
By relating it to the presence of extra matter, we can express it as a gauge-invariant local quantity, which is the natural thing to hold fixed in our analysis.
\end{enumerate}
Note that if we were reporting a positive result, that such uncertainty upholds unitarity, then one should question the validaty of learning a general lesson from a special example. However, we will show that this particular form of $\Delta g_{\mu\nu}$ is already sufficient to question unitarity, and we did not choose this form to specifically do that. Thus our limited analysis is sufficient to raise a reasonable doubt to the classical-background approximation.
\item {\bf Renormalization.} It is well-known that na\"ive applications of QFT often suffer from UV/IR divergences.
In our case, we will see that even in Minkowski space, geometric uncertainty already leads to a finite change between the two null geodesics.
We will simply ``renormalize'' that away by saying that if the black hole case leads to a similar change, it should not be taken as a serious problem.
Our surprising result is that in the black hole case, we get an infinitely larger change, which is difficult to blame on the usual divergence of QFT.
\end{itemize}

More concretely, our setup is shown in Fig.\ref{fig-setup}. 
We start by choosing two points at some $r_0\gg M$ with a $\delta \bar{t_0}\sim M$ coordinate time separation between them.
This represents one wavelength of the mode function of an asymptotic Hawking quantum.
We then back-track two null geodesics from these two points back to two points near the horizon, at $r_e$ with $(r_e-2M)\ll 2M$.
Now their separation represents one wavelength of a near-horizon mode, which would have propagated to the asymptotic mode if there were no geometric uncertainty. 
We then introduce introduce a geometric uncertainty as two shells of zero total ADM mass. 
Therefore, both the near horizon and the asymptotic regions are not affected by such uncertainty.
Only the process of propagation is affected.
With the same two starting points near the horizon we calculate the separation $\delta t_0$ of the two null geodesics when they arrive at $r_0$, and compare it with $\delta \bar{t_0}$.

\begin{figure}[tb]
\begin{center}
\includegraphics[scale = 0.6]{Propertime.pdf}
\caption{The two parts of this diagrams describe two different scenarios. The trajectories of the photon are the lines in orange. The trajectory of the observer and the position from which the photons are emitted are represented by solid blue lines. The bottom part is a general case of two Hawking photons coming from a distance $r_e$ from the black hole of mass $M$ and arriving at an observer who is at a distance of $r_0$. The top part shows two Hawking photons coming from the same distance $r_e$ from a black hole of mass $M$. Instead of the photons freely propagating through to an observer at $r_0$, they have to cross two shells, represented by blue dotted line, of equal and opposite mass $\Delta M, -\Delta M$ at $r_{1}, r_{2}$ respectively. The regions $I$ represent a metric with mass $M$ and region $II$ represents a metric with mass $M+\Delta M$.}
\label{fig-setup}
\end{center}
\end{figure}

@@

After some formal calculations in Appendix \ref{Dtime}, we arrive at this simple expression:
\begin{equation}
\frac{\delta t_0 - \bar{\delta t_0}}{\delta \bar{t_0}} = 
\left(\sqrt{\frac{r_2-2M}{r_1-2M}}\sqrt{\frac{r_1-2(M+\Delta M)}{r_2-2(M+\Delta M)}}-1\right)~,
\label{eq-result}
\end{equation}
where $r_1$ is the location of the inner shell, $r_2$ is the location of the outer shell, and $\pm\Delta M$ are their individual ADM masses.

@@

In the entire calculation, we will stay in the regime that $r_1, 2M, (r_1-2M) \gg \Delta M, \Delta r$, where $\Delta r \equiv(r_2-r_1)$. 
This guarantees that the geometric fluctuation is small, as it only affects a small region $\Delta r$, and the change in the metric $\Delta g_{\mu\nu}/g_{\mu\nu}$ is small in the affected region.
Under this assumption, we can expand Eq (\ref{eq-result}) to get 
\begin{equation}
	\frac{\delta t_0 - \bar{ \delta t_0}}{\delta \bar{t_0}} = \frac{\Delta M \Delta r}{(r_1-2M)^2} = \frac{\Delta M \Delta r}{r_1^2} g_{tt}^{-2}~.
		\label{lim-result}
\end{equation}
Here $g_{tt}=(1-2M/r_1)$ is the time component of the metric in the background geometry, evaluated at the location where the null rays pass through the geometric uncertainty.

It is more enlightening to write this expression in terms of local physical quantities (see Appendix \ref{finalEq} for derivations).  Let $\sigma$ be the energy density of the shell in its rest frame, and $\Delta l$ be the physical distance between the shells in that frame, we get
\begin{equation}
	\frac{\delta t_0 - \bar{ \delta t_0}}{\delta \bar{t_0}} = 4 \pi \sigma \Delta l g_{tt}^{-1}~.
	\label{eq-inv}
\end{equation}
Here we can see that the effect is due to the local energy density which the null rays pass through.
Also, the combination $(\sigma\Delta l)$ is invariant under radial boosts, thus Eq.~(\ref{eq-inv}) does not care about whether the shells are actually at rest, nor in which frame we calculate it.

@@

\begin{equation}
\frac{\delta t_0 - \bar{ \delta t_0}}{\delta \bar{t_0}}  \begin{cases}
= 4 \pi \sigma \Delta l~	 &\ \ \ \ \ (M = 0)\\
\approx 4\pi \sigma \Delta l \left( 1 + \frac{2M}{l} \right) & \ \ \ \ \ (M \ll l)	\\
\approx 4 \pi  \sigma \Delta l \frac{16M^2}{l^2} & \ \ \ \ \ (M \gg l).
\end{cases}	\label{fin-result}
\end{equation}

@@

Classically, soft hairs manifest as memory effects that alter the distances between geodesics, which is very similar to the effect of geometric uncertainty as we showed in this paper.
Two neighbouring null rays simply ``remember'' the geometric uncertainty they passed through.
Na\"ively speaking, a propagating mode of a massless field should remember a similar effect. 
Based on this picture, one may try to quantize the soft hairs and build a toy model that involves quantum interaction, therefore entanglement, between the soft hairs and the usual QFT modes.
That would provide a complete picture that the loss of subsystem unitarity is really just entanglement with another subsystem which was originally (and illegitimately, as we argued) assumed as the classical background.

@@

\begin{equation}
	ds^2 = - \left( 1 -\frac{2M}{r} \right) d\bar{t}^2 + \left( 1 - \frac{2M}{r} \right)^{-1} dr^2 + r^2 d \Omega_2^2	\label{1}
\end{equation}

where we are working with units in which $G = c = 1$. The bar is used over the time coordinate to differentiate it from the time coordinate we will use in the spacetime with mass $M'$. We know there will be a flux of particles appearing from the horizon of the black hole \cite{Haw74}. Take some coordinate time interval, $\delta \bar{t_e}$, (which will have some corresponding proper time interval $\delta \bar{\tau_e}$) between two consecutive particles being emitted from the same position above the horizon of the black hole. An observer far away will see these particles arrive with some coordinate time interval $\delta \bar{t_0}$. We can follow the trajectory of the particles (assuming they are relativistic) from the position they are emitted, $r_e$, to the observer at $r_0$ by setting $ds = 0$ in Eq (\ref{1}). By confining the motion to be only in the radial direction, i.e $d \Omega_2 = 0$, we get the following geodesic equation

\begin{equation}
	\bar{t_e}^{(1)} - \bar{t_0}^{(1)} = r_0 - r_e + 2M \ln \left( \frac{r_0 - 2M}{r_e - 2M}  \right)	\label{3}
\end{equation}

@@

\begin{equation}
	\bar{t_e}^{(2)} - \bar{t_0}^{(2)} = r_0 - r_e + 2M \ln \left( \frac{r_0 - 2M}{r_e - 2M}  \right).	\label{4}
\end{equation}

@@

Now we introduce two shells of matter with equal and opposite ADM mass, $\Delta M$ and $-\Delta M$ where $|\Delta M|<<M$, at a distance of $r_1$ and $r_2$ respectively as shown in the top part of figure \ref{fig-setup}. We model these shells as infinitesimally thin and therefore use the Israel Junction Conditions (IJC) \cite{Isr66} with a delta function shell to analyze the effects of the shells. In this case the particle's move through two different spacetime regions. Region $I$ is defined by the black hole of mass $M$ and region $II$ is defined by a Schwarzschild metric of mass $M' \equiv M+\Delta M$ as shown in the top part of figure \ref{fig-setup}. The metric in region II is given by

\begin{equation}
	ds^2 = - \left( 1 - \frac{2M'}{r} \right) d t^2 + \left( 1 - \frac{2M'}{r} \right)^{-1} dr^2 + r^2 d \Omega_2^2	\label{2}.
\end{equation}

Note that the time coordinate is different from the one in Eq (\ref{1}). The radial coordinate is however the same as the shell connecting the two spacetimes is a physical surface with a fixed radius $r$. As explained in the previous section, the proper time interval between two particles emitted at $r_e$, $\delta \bar{t_e}$, is the same as the coordinate time interval at $r_1$, $\delta \bar{t_1}$. Now we need to find the corresponding coordinate time interval in region $II$, $\delta t_1$. To do this we simply note that the corresponding proper time at $r_1$, $\delta \bar{\tau_1}$ must be the same in both regions of spacetime - this is a simple consequence of the IJC. The coordinate time intervals at $r_1$, in the two different regions, are therefore related by

\begin{equation}
	\delta \bar{t_1} = \delta \bar{t_e} = \left( \frac{1 - \frac{2(M + \Delta M)}{r_1}}{1 - \frac{2M}{r_1}} \right)^\frac{1}{2} \delta t_1.
\end{equation}

We can propagate the particles through region $II$ from $r_1$ to $r_2$ and we know $\delta t_1 = \delta t_2$. From the fact that the proper time corresponding to the coordinate time intervals will be the same, we can relate $\delta t_2$ back to $\delta \bar{t_2}$ in region $I$,
